---
title: "This is a test of Project 1"
subtitle: "Let's see how this goes." 
summary: "This is a summary of this post. Woo."
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Date: `r Sys.Date()`

Hello! This RMarkdown is for going through the first project of the book. :)

# R Interface 

The numbers before a result tell R what number result starts on that particular line. 
```{r}
100:130
```

Exercise 2.1, magic with numbers.You'll get the same number you start with by following that process.

```{r}
827 + 2

829 * 3

2487 - 6

2481 / 3
```

# Objects

Create a dice and store it in an object.
```{r}
die <- 1:6
```

See what objects are in your environment.
```{r}
ls()
```

R does what is called "element-wise execution" - this means that it does an operation to each element in a set. Here, we subtract one from each number on the die. 
```{r}
die - 1
```

This is essentially the principle behind tidy data - if you have one participant in each row, R will perfect operations within that row. 

You can get R to do matrix multiplication if you want it to. 
```{r}
# inner
die %*% die

# outer
die %o% die
```
# Functions

To randomly roll a die, you need to create a function. If you have multiple functions together, R works by solving the INNER most function first. In this example, the mean of the die is 3.5 so that is calculated first. Then that number is rounded to produce 4. 

```{r}
round(mean(die))
```

We can roll a die through an existing R function, `sample()`. 
```{r}
sample(x = die, size = 1)
```

Note that the `sample()` function has two arguments, `x` and `size`. To see how many arguments a function takes, you can execuute this:
```{r}
args(sample)
args(round)
```

Now let's simulate 2 rolls a die with replacement (the default is without, meaning you can't roll snake eyes)
```{r}
sample(x = die, size = 2, replace = TRUE)
```

# Writing your own functions

Writing a function called `roll()` that simulates rolling virtual dice. Each time you call it, R will return the sum of rolling two dice. Otherwise, you'd have to do it this way:
```{r}
die <- 1:6
dice <- sample(die, size = 2, replace = TRUE)
sum(dice)
```

3 basic parts: name, body of code, and set of arguments. Let's put the above code into a function:
```{r}
roll <- function() {
  die <- 1:6
  dice <- sample(die, size = 2, replace = TRUE)
  sum(dice)
}
```

Now let's run it:
```{r}
roll()
```

Typing the function WITHOUT parentheses will give you what is contained insidie the function:

```{r}
roll
```
Remember that R will execute ALL of the code int he body and the result of the LAST line of code. If your function isn't written in a way where that will happen, then you won't get a result. 

## Arguments

Now let's create an argument called `bones` where you can specify different types of die (1:6 or 1:20), and let's create that so the default is 1:6. 
```{r}
roll2 <- function(bones = 1:6) {
  dice <- sample(bones, size = 2, replace = TRUE)
  sum(dice)
}
```

Now run the default:
```{r}
roll2()
```

Now 20-sided dice:
```{r}
roll2(bones = 1:20)
```

One thing that's really cool about this is that writing your own functions is basically like adding in your own command to a program like Excel or SPSS. 

# Packages and Help Pages

You know how to work with packages so there's no real need to document this section :) 

Let's make our first `qplot` (quick plot) with ggplot 2.
```{r, warning = F}
library(ggplot2)

x <- c(-1, -0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8, 1)

y <- x^3

qplot(x, y)
```

Now let's look at a histogram to inspect one vector (i.e., variable). 
```{r, warning = F}
x <- c(1, 2, 2, 2, 3, 3)
qplot(x, binwidth = 1)
```

And now another one:
```{r, warning = F}
x2 <- c(1, 1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 4)
qplot(x2, binwidth = 1)
```

Exercise 3.1 (visualize a histogram). 
```{r, warning = F}
x3 <- c(0, 1, 1, 2, 2, 2, 3, 3, 4)
qplot(x3, binwidth = 1)
```

Ok so now we're going to see how to repeat our dice rolling function many times. This can be done with the `replicate()` function. Replicate runs the same command many times and then stores the results as a vector. 
```{r}
replicate(3, 1 + 1)

replicate(10, roll())
```

Let's go ahead and simualte 10,000 dice rolls to see what this will look like.
```{r, warning = F}
rolls <- replicate(10000, roll())
qplot(rolls, binwidth = 1)
```

This is symmetric, as we'd expect! This is proportional to the odds of the various combinations of dice. This corresponds with a 1/6 probability of any number getting rolled. What if we increased getting a 6 to 3/8 (and all others to 1/8)? 

```{r}
roll <- function() {
  die <- 1:6
  dice <- sample(die, size = 2, replace = TRUE, prob = c(1/8, 1/8, 1/8, 1/8, 1/8, 3/8))
  sum(dice)
}
roll()
```

```{r}
rolls <- replicate(10000, roll())
qplot(rolls, binwidth = 1)
```

Et voilà, we created an unfair set of dice here :D 
